---
title: "clean_raw_expedata"
author: "Dani Blumstein"
date: "12/2/2020"
output: html_document
---

bring in libs
```{r}
library(devtools)
library(tidyverse)
library(lubridate)
library(readr)
library(viridis)
library(patchwork)
library(tidyselect)
library(readxl)
```

Important equations:
RQ = CO2 eliminated/O2 consumed
EE = 0.06 * (3.941 * VO2 + 1.106 * VCO2)

from 14.4 in Leighton book
LabDiet 5015 = (26.101/100).71+(19.752/100).83+.54148 = .8907387
LabDiet low fat 5015 = (22.8/100).71+(6.6/100).83+.706 = .92266

Functions:
import data function
```{r}
bring_in_data <- function(data_file, Sex)
{
  data <- paste(path,data_file,sep="")
  raw <- read_csv(data,
        col_types = cols(Animal = col_double(), 
        StartDate = col_date(format = "%m/%d/%Y"),
        deltaCO2 = col_double(), 
        deltaH2O = col_double(),
        H2Oml = col_double(),
        Deg_C = col_double(),
        VCO2 = col_double(),
        StartTime = col_time(format = "%H:%M:%S")))
  
  '%!in%' <- function(x,y)!('%in%'(x,y))

raw <- raw %>% 
	mutate(EE = 0.06*(3.941*VO2 + 1.106*VCO2)) %>% 
	mutate(RQ = VCO2/VO2) %>%
	mutate(animal = round(Animal, digits=0)) %>%
	mutate(Animal = NULL) %>%
  mutate(Sex  = Sex) %>%
	unite("DateTime", StartDate:StartTime, remove = FALSE, sep =  " ") %>%
	mutate(weight = 
		ifelse(animal == 0, cageweight0, 
		ifelse(animal == 1, cageweight1,
		ifelse(animal == 2, cageweight2,
		ifelse(animal == 3, cageweight3,
		ifelse(animal == 4, cageweight4,
		ifelse(animal == 5, cageweight5,
		ifelse(animal == 6, cageweight6, NA)))))))) %>% 
	mutate(Animal_ID = 
		ifelse(animal == 0, animalID0, 
		ifelse(animal == 1, animalID1,
		ifelse(animal == 2, animalID2,
		ifelse(animal == 3, animalID3,
		ifelse(animal == 4, animalID4,
		ifelse(animal == 5, animalID5,
		ifelse(animal == 6, animalID6, NA)))))))) %>% 
	mutate(H2Omg_edit = 
		ifelse(hour(StartTime) == 8, H2Omg,
		ifelse(hour(StartTime) == 7, H2Omg,
		ifelse(hour(StartTime) == 9, H2Omg,
		ifelse(hour(StartTime) == 10, H2Omg,
		ifelse(hour(StartTime) == 19, H2Omg,
		ifelse(hour(StartTime) == 20, H2Omg,
		ifelse(hour(StartTime) == 21, H2Omg,
		ifelse(hour(StartTime) == 22, H2Omg,
		ifelse(hour(StartTime) %!in% c(7,8,9,10,20,21,22,19), H2Omg, NA)))))))))) %>% 
	mutate_at("H2Omg_edit", as.numeric)

#metric <- "corEE"

target <- c(0,1,2,3,4,5,6,7)
cages <- raw %>% filter(animal %in% target)

  
  #start_time <- ymd_hms(subset[[5]][1])
  #begin_experiment <- start_time + dhours(2)
  #end_time <- begin_experiment + dhours(72)
  #filtered <- subset %>% filter(raw$DateTime >= begin_experiment & raw$DateTime <= end_time)
  
  
  return(cages)
}
```

mouse id and weights function
will add electrolytes....one day
```{r}
weight_ID <- function(date)
{  
  subset <- electrolyte_data[which(electrolyte_data$experiment_date == date), names(electrolyte_data) %in% c("sex", "mouse_ID", "cage_number", "weight", "Na", "K", "Cl", "TCO2", "BUN", "Crea", "Glu", "iCa", "AnGap", "Hct", "Hb*")]
  
  ids <- subset$mouse_ID
  weights <- as.double(subset$weight)
  
  x = 0

  for (i in 1:length(ids))
  {
    assign(paste("animalID", x, sep = ""), ids[i], envir = parent.frame())
    assign(paste("cageweight", x, sep = ""), weights[i], envir = parent.frame())
    x = x + 1
  }
} 
```

merge data and subset for 72 hours function
```{r}
merge_data <- function(cage)
{
  start_time <- ymd_hms(cage[[1]][1])
  begin_experiment <- start_time + dhours(2)
  end_time <- begin_experiment + dhours(72)
  filtered <- cage %>% filter(DateTime >= begin_experiment & DateTime <= end_time)
  
  return(filtered)
}
```

import data
electrolyte and weight data
and path to files. this should be one of the only things edited for now 
```{r}
path <- "~/Box Sync/Cactus_Mouse_Physiology/data/"
electrolyte_data <- read_excel(paste(path,"electrolyte_data.xlsx",sep=""), na = "NA")
```

males 1 baseline
Feb 26, 2020
```{r}
weight_ID("26-Feb-20")
cages26feb <- merge_data(bring_in_data("baseline/26Feb20/feb26.csv", "M"))
```

males 2 baseline
Mar 10, 2020
```{r}
weight_ID("5-Mar-20")
cages10mar <- merge_data(bring_in_data("baseline/10Mar20/mar10.csv", "M"))
```

females 1 baseline
Feb 20, 2020
```{r}
weight_ID("20-Feb-20")
cages20feb <- merge_data(bring_in_data("baseline/20Feb20/feb20.csv", "F"))
```

females 2 baseline
March 14, 2020
```{r}
weight_ID("14-Mar-20")
cages14mar <- merge_data(bring_in_data("baseline/14Mar20/mar14.csv", "F"))
```

females1 constant 70 deg temp
April 20, 2020
```{r}
weight_ID("20-Apr-20")
cages20april1 <- bring_in_data("cold/20April20/20_april_2020_1.csv", "F")

weight_ID("20-Apr-20")
cages20april2 <- bring_in_data("cold/20April20/20_april_2020_2.csv", "F")

weight_ID("20-Apr-20")
cages20april3 <- bring_in_data("cold/20April20/20_april_2020_3.csv", "F")

weight_ID("20-Apr-20")
cages20april4 <- bring_in_data("cold/20April20/20_april_2020_4.csv", "F")

cages20april_all <- rbind(cages20april1, cages20april2, cages20april3, cages20april4)

cages20april <- merge_data(rbind(cages20april2, cages20april3, cages20april4))
```

males1 constant 70 deg temp
April 27, 2020
```{r}
weight_ID("27-Apr-20")
cages27april1 <- bring_in_data("cold/27April20/27_april_2020_1.csv", "M")

weight_ID("27-Apr-20")
cages27april2 <- bring_in_data("cold/27April20/27_april_2020_2.csv", "M")

weight_ID("27-Apr-20")
cages27april3 <- bring_in_data("cold/27April20/27_april_2020_3.csv", "M")

cages27april <- merge_data(rbind(cages27april1, cages27april2, cages27april3))
```

females2 constant 70 deg temp
April 30, 2020
```{r}
weight_ID("30-Apr-20")
cages30april1 <- bring_in_data("cold/30April20/30_april_2020_1.csv", "F")

weight_ID("30-Apr-20")
cages30april2 <- bring_in_data("cold/30April20/30_april_2020_2_wrong_temp.csv", "F")

weight_ID("30-Apr-20")
cages30april3 <- bring_in_data("cold/30April20/30_april_2020_3_wrong_temp.csv", "F")

weight_ID("30-Apr-20")
cages30april4 <- bring_in_data("cold/30April20/30_april_2020_4.csv", "F")

weight_ID("30-Apr-20")
cages30april5 <- bring_in_data("cold/30April20/30_april_2020_5.csv", "F")

weight_ID("30-Apr-20")
cages30april6 <- bring_in_data("cold/30April20/30_april_2020_6.csv", "F")

cages30april_all <- rbind(cages30april1, cages30april2, cages30april3, cages30april4, cages30april5, cages30april6)

cages30april <- merge_data(rbind(cages30april4, cages30april5, cages30april6))
```

males 2 constant 90 deg temp
May 18, 2020
```{r}
weight_ID("18-May-20")
cages18may1 <- bring_in_data("hot/18May20/18_may_20_1.csv", "M")

weight_ID("18-May-20")
cages18may2 <- bring_in_data("hot/18May20/18_may_20_2.csv", "M")

weight_ID("18-May-20")
cages18may3 <- bring_in_data("hot/18May20/18_may_20_3.csv", "M")

weight_ID("18-May-20")
cages18may4 <- bring_in_data("hot/18May20/18_may_20_4.csv", "M")

cages18may_all <- rbind(cages18may1, cages18may2, cages18may3, cages18may4)
cages18may <- merge_data(rbind(cages18may2, cages18may3, cages18may4))
```

Females 1 constant 90 deg temp
May 25, 2020
```{r}
weight_ID("25-May-20")
cages25may1 <- bring_in_data("hot/25May20/25_may_20_1.csv", "F")

weight_ID("25-May-20")
cages25may2 <- bring_in_data("hot/25May20/25_may_20_2.csv", "F")

weight_ID("25-May-20")
cages25may3 <- bring_in_data("hot/25May20/25_may_20_3.csv", "F")

weight_ID("25-May-20")
cages25may4 <- bring_in_data("hot/25May20/25_may_20_4.csv", "F")

cages25may <- merge_data(rbind(cages25may1, cages25may2, cages25may3, cages25may4))
```

Females 2 constant 90 deg temp
May 29, 2020
```{r}
weight_ID("29-May-20")
cages29may1 <- bring_in_data("hot/29May20/29_may_20_1.csv", "F")

weight_ID("29-May-20")
cages29may2 <- bring_in_data("hot/29May20/29_may_20_2.csv", "F")

weight_ID("29-May-20")
cages29may3 <- bring_in_data("hot/29May20/29_may_20_3.csv", "F")

cages29may <- merge_data(rbind(cages29may1, cages29may2, cages29may3))
```

Males 1 constant 90 deg temp
June 2, 2020
```{r}
weight_ID("2-Jun-20")
cages3june1 <- bring_in_data("hot/3June20/3_june_20_1.csv", "M")

weight_ID("2-Jun-20")
cages3june2 <- bring_in_data("hot/3June20/3_june_20_2.csv", "M")

weight_ID("2-Jun-20")
cages3june3 <- bring_in_data("hot/3June20/3_june_20_3.csv", "M")

weight_ID("2-Jun-20")
cages3june4 <- bring_in_data("hot/3June20/3_june_20_4.csv", "M")

cages3june <- merge_data(rbind(cages3june1, cages3june2, cages3june3, cages3june4))
```

males2 constant 70 deg temp
June 7, 2020
```{r}
weight_ID("7-Jun-20")
cages7june1 <- bring_in_data("cold/7June20/7_june_20_1.csv", "M")

weight_ID("7-Jun-20")
cages7june2 <- bring_in_data("cold/7June20/7_june_20_2.csv", "M")

weight_ID("7-Jun-20")
cages7june3 <- bring_in_data("cold/7June20/7_june_20_3.csv", "M")

weight_ID("7-Jun-20")
cages7june4 <- bring_in_data("cold/7June20/7_june_20_4.csv", "M")

cages7june <- merge_data(rbind(cages7june2, cages7june3, cages7june4))
```

Establish when each interval/transition starts and stops in seconds
```{r}
#Daytime interval: hrs:8:00-21:00
daytime_interval <- period_to_seconds(hms("09:00:00")):period_to_seconds(hms("20:00:00"))
#Night time: hrs 22:00-5:00 (do NOT do a 22:5 sequence as it will count backwards from 22 to 5...)
nighttime_interval <- c((period_to_seconds(hms("21:00:01")):period_to_seconds(hms("24:59:59"))), #evening portion of 'nighttime'
                        (period_to_seconds(hms("00:00:00")):period_to_seconds(hms("06:00:00")))) #morning portion of 'nightitme'

```

"final" data sets
```{r}
baseline <- rbind(cages26feb, cages10mar, cages20feb, cages14mar)
baseline$experiment <- "baseline"
baseline$FQ <- 0.8907387
baseline$seconds <- as.numeric(baseline$StartTime)
#baseline$TimeOfDay <- ifelse(baseline$seconds %in% nighttime_interval, "night", "day")
target <- c(0,1,2,3,4,5,6)
baseline <- baseline %>% filter(animal %in% target)

cold <- rbind(cages20april, cages27april, cages30april, cages7june)
cold$experiment <- "cold"
cold$FQ <- 0.8907387
cold$seconds <- as.numeric(cold$StartTime)
#cold$TimeOfDay <- ifelse(cold$seconds %in% nighttime_interval, "night", "day")
target <- c(0,1,2,3,4,5,6)
cold <- cold %>% filter(animal %in% target)

hot <- rbind(cages18may, cages25may, cages29may, cages3june)
hot$experiment <- "hot"
hot$FQ <- 0.8907387
hot$seconds <- as.numeric(hot$StartTime)
#hot$TimeOfDay <- ifelse(hot$seconds %in% nighttime_interval, "night", "day")
target <- c(0,1,2,3,4,5,6)
hot <- hot %>% filter(animal %in% target)
```

select your data set for all down stream code. this should be changed depending on what down stream analysis you are doing 
```{r}
all_animals <- rbind(hot, cold, baseline)

length(all_animals)

#remove empty cage 
target <- c(0,1,2,3,4,5,6)
cages <- all_animals %>% filter(animal %in% target)
```

getting data clean and analysis ready 
```{r}
analysis_data <- cages[,c("Sex","EE","H2Omg", "RQ", "StartTime","Animal_ID", "VO2", "VCO2", "Deg_C", "weight", "experiment", "SD_VCO2", "SD_VO2", "StartDate", "SD_H2Omg", "FQ")]

#split the time up by hour to make it an easier vairable to work with
analysis_data_edited <- data.frame(do.call('rbind', strsplit(as.character(analysis_data$StartTime),':',fixed=TRUE)))

#add it back to the subseted data for analysis
analysis_data_inter <- cbind(analysis_data,analysis_data_edited)
analysis_data_final <- analysis_data_inter[,c("Sex","EE","H2Omg", "RQ","Animal_ID", "Deg_C", "weight", "experiment", "X1", "StartTime", "SD_VCO2", "SD_VO2", "SD_H2Omg", "VO2", "VCO2", "StartDate", "FQ")]

analysis_data_final$Sex <- as.factor(analysis_data_final$Sex)
analysis_data_final$hour <- analysis_data_final$X1
analysis_data_final$X1 <- NULL
analysis_data_final$Deg_C <- as.double(analysis_data_final$Deg_C)
analysis_data_final$Animal_ID <- as.numeric(analysis_data_final$Animal_ID)
analysis_data_final$experiment <- as.factor(analysis_data_final$experiment)

names(analysis_data_final)[3] <- "H2Omg"


unique(analysis_data_final$Animal_ID)
write_csv(analysis_data_final, path = "/Users/danielleblumstein/Desktop/analysis_data_final.csv")
```


